---
title: "Burrito Exercise"
author: "Sydney Dieter,  Wulf Novak,  Xie Shuchen"
date: "STAT 440 | Spring 2018"
output:
  ioslides_presentation:
    widescreen: yes
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10)
options(width = 65, digits=3)
```

## Research Question



- How does hunger level and cost effect the overall rating of the burrito? 

## Data processing

- We prepared the data by removing the rows that didn't have a data entry. 



```{r,echo=FALSE}
 rm(list=ls())
 load("burritodata.Rda")
 data = burrito
 Dat = cbind(data$overall, data$Cost, data$Hunger, data$Reviewer)
 x = Dat[, -1] 
 y = Dat[, 1] 
```

```{r,results='hide'}
 stopifnot(complete.cases(y) != is.na(y))
 ok = complete.cases(x, y)
 x = x[ok,]
 y = y[ok]
 z = x[,3]
``` 

```{r,echo=FALSE}
x <- data.matrix(cbind(as.numeric(x[,1]),as.numeric(x[,2])))
y <- as.numeric(y)
```

## Model
1) BLMM 
2) BLM

Block Gibbs sampler 

$$Y_{ij}|\alpha, \beta, \gamma_j, \sigma^{-2} \sim N(\alpha + X_{ij}^T\beta + \gamma_j, \sigma^2)$$
$$\alpha \sim Flat$$
$$\beta \sim N(0, \tau^2I)$$
$$\gamma_j|\kappa^{-2} \sim N(0, \kappa^2)$$
$$\sigma^{-2} \sim Gamma(a_1, a_2)$$
$$\kappa^{-2} \sim Gamma(b_1, b_2)$$



## Diagnostics

- Check convergence

```{r,echo=FALSE}
BLMM <- function(y,x,z,S=10000,mu=0,tau=1000,a1=0.01, a2=0.01, b1=0.01, b2=0.01){

 # create the variables now to avoid repeating this calculation each step
 n <- nrow(x)
 p <- ncol(x)
 if(is.null(p)) p <- 1
 g <- length(unique(z))
 X <- cbind(1,x,model.matrix(~z-1)) # add intercept and random effect indicators
 Xy <- t(X)%*%y
 XX <- t(X)%*%X

 # Places to store results
 beta_results <- matrix(NA, S, ncol(X))
 sig2inv_results <- rep(NA,S)
 kappa2inv_results <- rep(NA,S)

 # starting values
 sig2inv <- 1/var(y)
 kappa2inv <- 1/var(aggregate(y,by=list(z),mean)$x)

 # MCMC!!!!
 for(s in 1:S){

 # update beta including intercept
 tau2inv <- c(0,rep(1/tau^2,p), rep(kappa2inv,g))
 v <- chol2inv(chol((sig2inv*XX+diag(tau2inv))))
 beta <- t(chol(v)) %*% rnorm(ncol(X)) + v %*% ( sig2inv*Xy +tau2inv*mu )

 #update sig2inv
 sig2inv <- rgamma(1, a1 + length(y)/2, a2 + sum((y-X%*%beta)^2)/2 )

 #update sig2inv
 kappa2inv <- rgamma(1, b1 + g/2, b2 + sum((beta[-c(1:(1+p))])^2)/2 )
 # beta[2:3]

 # store results
 beta_results[s,] <- beta
 sig2inv_results[s] <- sig2inv
 kappa2inv_results[s] <- kappa2inv

 }



 return(list(alpha=beta_results[,1],beta1=beta_results[,2],beta2=beta_results[,3],
 gamma=beta_results[,-c(1:(1+p))],kappa=sqrt(1/kappa2inv_results),sigma=(sqrt(1/sig2inv_results))))



 }
 Model1 = BLMM(y,x,z)
 
 par(mfrow=c(2,3))
 # par(mfrow=c(1,1))
 plot(Model1$alpha, type="l", las=1)
 plot(Model1$beta1, type="l", las=1)
 plot(Model1$beta2, type="l", las=1)
 matplot(Model1$gamma, type="l", las=1)
 plot(Model1$kappa, type="l", las=1)
 plot(Model1$sigma, type="l", las=1)
```

## Diagnostics

- ACF

```{r,echo=FALSE}
 par(mfrow=c(1,3))
 acf(Model1$alpha)
 acf(Model1$beta1)
 acf(Model1$beta2)
```

## Diagnostics

- Check convergence

```{r,echo=FALSE}
BGS = function(y,x,S=10000, a1=0.01, a2=0.01){

 n<- length(y)
 X <- data.matrix(cbind(rep(1,n), x))
 XX <-t(X)%*%X
 XY<-t(X)%*%y


 mu = data.matrix(c(0,0,0))
 sig2inv=1
 tau2inv = data.matrix(cbind(c(0,0,0),c(0,1/1000^2,0),c(0,0,1/1000^2)))

 beta_0_samples = numeric(S)
 beta_1_samples = numeric(S)
 beta_2_samples = numeric(S)
 sigma_2_inv_samples = numeric(S)

 for (s in 1:S){
 v<- chol2inv(chol((sig2inv*XX+diag(tau2inv))))
 beta<- t(chol(v))%*%rnorm(ncol(X))+v%*%(sig2inv*XY+tau2inv%*%mu)
 sig2inv<-rgamma(1,a1+length(y)/2,a2+sum((y-X%*%beta)^2)/2)

 beta_0_samples[s] = beta[1,1]
 beta_1_samples[s] = beta[2,1]
 beta_2_samples[s] = beta[3,1]
 sigma_2_inv_samples[s] = sig2inv
 }
 
 sigma_samples = 1/sqrt(sigma_2_inv_samples)

 return(list(beta_0=beta_0_samples,beta_1=beta_1_samples,beta_2=beta_2_samples,sigma=sigma_samples))
 }

 Model2 = BGS(y,x)
 
 par(mfrow=c(2,2))
 plot(Model2$beta_0, type="l", las=1)
 plot(Model2$beta_1, type="l", las=1)
 plot(Model2$beta_2, type="l", las=1)
 plot(Model2$sigma, type="l", las=1)
```

## Diagnostics

- ACF

```{r,echo=FALSE}
 par(mfrow=c(1,3))
 acf(Model2$beta_0)
 acf(Model2$beta_1)
 acf(Model2$beta_2)

```


## Results

```{r,echo=FALSE}
 results1 <- data.frame(
 mean = c(mean(Model1$alpha), mean(Model1$beta1),
 mean(Model1$beta2), mean(Model1$kappa), 
 mean(Model1$sigma)),
 sd = c(sd(Model1$alpha), sd(Model1$beta1),
 sd(Model1$beta2), sd(Model1$kappa), 
 sd(Model1$sigma)),
 quantile_lower = c(quantile(Model1$alpha, 0.025), quantile(Model1$beta1, 0.025),
 quantile(Model1$beta2, 0.025), quantile(Model1$kappa, 0.025),
 quantile(Model1$sigma, 0.025)),
 quantile_upper = c(quantile(Model1$alpha, 0.975), quantile(Model1$beta1, 0.975),
 quantile(Model1$beta2, 0.975), quantile(Model1$kappa, 0.975),
 quantile(Model1$sigma, 0.975)) 
 )
 row.names(results1) <- c("alpha", "beta1", "beta2","kappa", "sigma")
 results1

```



```{r,echo=FALSE}
 results2 <- data.frame(
 mean = c(mean(Model2$beta_0), 
 mean(Model2$beta_1),
 mean(Model2$beta_2),
 mean(Model2$sigma)),
 sd = c(sd(Model2$beta_0), 
 sd(Model2$beta_1),
 sd(Model2$beta_2),
 sd(Model2$sigma)),
 quantile_lower = c(quantile(Model2$beta_0, 0.025), 
 quantile(Model2$beta_1, 0.025),
 quantile(Model2$beta_2, 0.025), 
 quantile(Model2$sigma, 0.025)),
 quantile_upper = c(quantile(Model2$beta_0, 0.975),
 quantile(Model2$beta_1, 0.975),
 quantile(Model2$beta_2, 0.975), 
 quantile(Model2$sigma, 0.975)) 
 )
 row.names(results2) <- c("beta0", "beta1", "beta2", "sigma")
 results2

```

## Conclusions

 
The reviewer did not greatly effect the outcome of the overall rating of the burrito. We found that the hunger level was over twice as important as the cost of the burrito. 
